---
title: "HW: tidyr pivots"
output: html_notebook
---

---------------

Objective: Practice tidyr pivot skills

Task: Follow prompts below. Enter code in code chunks.

Help: [tidyr Documentation]("https://tidyr.tidyverse.org/")

Submission: Submit on Canvas. Remember to change file name to include your name as templated.

Total points: 16

------------------

**Question 1 (1 point)**   
Import the `tidyverse` package

```{r}
# Import libraries
library(tidyverse)
```

**Question 2 (1 point)**   
Load the `example_eagle_pairs.csv`   
This .csv contains data on the number of observed bald eagle breeding pairs across years and states.
```{r}
# Code
eagles <- read_csv("example_eagle_pairs.csv")
eagles
```

**Question 3 (1 point)**   
Rename column headers to remove `X`
```{r}
# Code
#I went to OH and talked to professor Kupperman about the data set. Once I
#uploaded the dataset the `X` was already removed from the columns. 
#I am writing the potential code for the problem if I had `X` . 


#eagles <- rename(eagles, `1997` = X1997, `1998` = X1998, `1999` = X1999, `2000` = X2000, `2001` = X2001, `2002` = X2002, `2003` = X2003, `2004` = X2004, `2005` = X2005, `2006` = X2006)

```

**Question 4 (2 points)**  
List the variables contained in dataset
```{r}
# Answer
str(eagles)
```

**Question 5 (1 point)**
In one sentence explain why this data isn't tidy

```{r}
# Answer
#There are variables (year and number of nests) that are not in a singular columns
```

**Question 6 (3 points)**   
Pivot this dataset longer. 
```{r}
# Code
eagles %>% 
  pivot_longer(
    cols = c(`1997`:`2006`), 
    names_to = "year", 
    values_to = "num_nests"
  )

```

**Question 7 (1 point)**   
Import `example_acs_1.csv`

```{r}
# Code
example_acs_1 <- read_csv("example_acs_1.csv")
View(example_acs_1)
```

**Question 8 (1 point)**
This is an example of a dataset that need to pivoted wider in order to be tidy. Which 2 variables are listed as variables but really not?

```{r}
# Answer
#GeoID and Variable 
```

**Question 9 (1 point)**
How many distinct variables are listed in `variable`? (use dplyr verbs)

```{r}
# Code
example_acs_1 %>% distinct(example_acs_1$variable)

# Answer
# There are 3 listed in `variable`,  pop_housed, pop_renter, median_renter
```

**Question 10 (1 point)**
For the distinct values listed in Question 9, in what column are their values currently stored?

```{r}
# Answer
#Their values are currently stored in the column 'variable'
```

**Question 11 (2 points)**
Use your answers from above to pivot this dataset wider. 

```{r}
# Code
example_acs_1 %>% 
  pivot_wider(
    names_from = variable,
    values_from = estimate
    ) 

```

**1 point: readable, tidyverse code**